---
format: gfm
knitr: 
  opts_chunk: 
    collapse: true
    comment: "#>"
    fig.path: "man/figures/README-"
    out.width: "100%"
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

# tidynorm <a href="https://jofrhwld.github.io/tidynorm/"><img src="man/figures/logo.svg" alt="tidynorm website" align="right" height="139"/></a>

<!-- badges: start -->

[![R-CMD-check](https://github.com/JoFrhwld/tidynorm/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/JoFrhwld/tidynorm/actions/workflows/R-CMD-check.yaml)

<!-- badges: end -->

The goal of `{tidynorm}` is to provide convenient and tidy functions to normalize vowel formant data.

## Installation

You can install the development version of tidynorm like so:

``` r
## if you need to install `remotes`
# install.packages("remotes")
remotes::install_github("jofrhwld/tidynorm")
```

## Example

Vowel formant frequencies are heavily influenced by vocal tract length differences between speakers. Equivalent vowels between speakers can have dramatically different frequency locations.

```{r}
#| label: unnorm
library(tidynorm)
library(ggplot2)
options(
  ggplot2.discrete.colour = \(...) scale_color_brewer(palette = "Dark2", ...)
)

ggplot(
  speaker_data,
  aes(
    F2, F1,
    color = speaker
  )
)+
  stat_density_2d(
    bins = 4
  )+
  scale_x_reverse()+
  scale_y_reverse()+
  coord_fixed()+
  labs(
    title = "unnormalized"
  )
```

The goal of `{tidynorm}` is to provide tidyverse-friendly and familiar functions that will allow you to quickly normalize vowel formant data. There are a number of built in functions based on conventional normalization methods.

```{r}
#| label: norm
speaker_data |> 
  norm_nearey(
    F1:F3,
    .by = speaker,
    .names = "{.formant}_nearey"
  ) ->
  speaker_normalized

speaker_normalized |> 
  ggplot(
    aes(
      F2_nearey, F1_nearey,
      color = speaker
    )
  )+
  stat_density_2d(
    bins = 4
  )+
  scale_x_reverse()+
  scale_y_reverse()+
  coord_fixed() +
  labs(
    title = "Nearey Normalized"
  )
```

There is also a `tidynorm::norm_generic()` function to allow you to define your own bespoke normalization methods. For example, a "robust Nearey" normalization method using the median, instead of the mean, could be done like so.

```{r}
speaker_rnearey <- speaker_data |> 
  norm_generic(
    F1:F3,
    .by = speaker,
    .by_formant = FALSE,
    .pre_trans = log,
    .L = median(.formant, na.rm = T),
    .names = "{.formant}_rnearey"
  )

speaker_rnearey |> 
 ggplot(
    aes(
      F2_rnearey, F1_rnearey,
      color = speaker
    )
  )+
  stat_density_2d(
    bins = 4
  )+
  scale_x_reverse()+
  scale_y_reverse()+
  coord_fixed() +
  labs(
    title = "Robust Nearey Normalized"
  )
```
